---
title: "US Border Apprehensions Data Analysis"
author: "Nanfang Hong, Jiajian Guo"
date: "2/9/2018"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Background

**US-Mexico border apprehensions hit 17-year lows**

(Washington, CNN) Apprehensions at the US-Mexico border reached historic lows in April, continuing a downward slide in the first few months of the Trump administration.

## Data Visualization

```{r, echo = FALSE}
df2010 <- read.csv("BP Apprehensions 2010.csv")
dMat2010 <- gsub(",", "", as.matrix(df2010[1:9,2:13]))
storage.mode(dMat2010) <- "numeric"
df2017 <- read.csv("PB Apprehensions 2017.csv")
dMat2017 <- gsub(",", "", as.matrix(df2017[1:9,2:13]))
storage.mode(dMat2017) <- "numeric"

monthMat <- rbind(colSums(dMat2010), colSums(dMat2017))
rownames(monthMat) <- c(2010, 2017)

sectorMat <- rbind(rowSums(dMat2010), rowSums(dMat2017))
colnames(sectorMat) <- df2010$Sector
rownames(sectorMat) <- c(2010, 2017)
```


```{r}

barplot(monthMat, main = "2010 vs. 2017 Comparison of Apprehension by month",
  las=2, col=c("darkblue","red"),
 	legend = rownames(monthMat), beside=TRUE)
```

## Data Visualization

```{r}
#opts_chunk$set(dev = 'pdf')
barplot(sectorMat, main = "2010 vs. 2017 Comparison of Apprehension by sector",
  las=2, col=c("yellow","green"),
 	legend = rownames(sectorMat), beside=TRUE)
```

## Questions

- Has there been a change in the sector with the most apprehensions for 2010 and with the sector with the most apprehensions in 2017?
- 3 month periods with the most apprehensions in 2010 and 2017?

## Statistical Analysis

- The sector with the most apprehensions for 2010 is Tucson. 
- The sector with the most apprehensions for 2017 is Rio Grande Valley. 
- Then we do a two-sample Welch t-test. As a result, the p-value is 0.06346, showing no significant change in the maximum. 

## Statistical Analysis
- 3 month period for 2010 and the most is March-April-May
- 3 month period for 2017 and the most is October-November-December
- Then we do a two-sample Welch t-test again on 3 month periods with the most apprehensions in 2010 and 2017. As a result, the p-value is 0.4042, showing no significant change in the maximum 3 consecutive monthly apprehensions. 

## Historical Data

```{r}
# Usual month sequence
ts2 <- read.csv("PB monthly summaries.csv")
#ts3 <- cbind(ts2[5:13], ts2[2:4])
#ts1 <- rev(as.vector(t(rev(ts3))))
#ts1 <- ts(ts1, start = c(2000,10), frequency=12)
#ts.plot(ts1, gpars=list(xlab="year", ylab="Apprehensions", lty=c(1:3)))

#Last month of October 2017 is September 2016
ts4 <- ts2[2:13]
ts5 <- rev(as.vector(t(rev(ts4))))
ts5 <- ts(ts5, start = c(2000,10), frequency=12)
ts.plot(ts5, gpars=list(xlab="year", ylab="Apprehensions", lty=c(1:3)), main = "Historical Apprehension data from 2000 to 2017")


ts7 <- as.matrix(ts4)
ts8 <- rev(sapply(1:18, function(i) sum(ts7[i, ]) / 12))
namebank <- as.character(c(2000:2017))
text(c(2000:2017) + 1, ts8, labels=namebank, cex= 0.7)
text(c(2000:2017) + 1, ts8, labels="----", cex= 0.7)
```

